---
title: "Adding arrows to confidence intervals when using shape_plot()"
date: "2022-11-05"
author: "Neil Wright"
---
```{r}
#| include: false
library(ckbplotr)
```
Using `ckbplotr::shape_plot()` you may have confidence interval limits outside your chosen limits for the y axis. The confidence interval lines (and text labels) will extend outside the plotting area and maybe outside the plot entirely:
```{r}
#| warning: false
shape_plot(ckbplotr_shape_data[ckbplotr_shape_data$is_female == 0,],
           col.x        = "rf",
           col.estimate = "est",
           col.stderr   = "se",
           col.n        = "n",
           xlims        = c(15, 50),
           ylims        = c(0.9, 1.8),
           scalepoints  = TRUE)
```

If you definitely want to keep those axis limits, you may want to add arrows to the confidence interval lines. First use `addaes` in `ckbplotr::shape_plot()` to truncate the confidence interval lines (and change the position of the text):
```{r}
#| warning: false
p <- shape_plot(ckbplotr_shape_data[ckbplotr_shape_data$is_female == 0,],
                col.x        = "rf",
                col.estimate = "est",
                col.stderr   = "se",
                col.n        = "n",
                xlims        = c(15, 50),
                ylims        = c(0.9, 1.8),
                scalepoints  = TRUE,
                printplot    = FALSE,
                addaes       = list(ci = c("ymin = pmax(0.9, est-1.96*se)",
                                           "ymax = pmin(1.8, est+1.96*se)"),
                                    estimates = "y = pmin(1.8, est+1.96*se)",
                                    n         = "y = pmax(0.9, est-1.96*se)"),
                
                ## increase the gap between the x-axis and plotting area
                gap = c(0.025, 0.1))
```

Then add `geom_segment()` layers to the plot to add small segments with arrows. Use the `data` argument to filter the data to only include the rows where the confidence limit is outside the axis limit.
```{r}
#| warning: false
p$plot +
  geom_segment(aes(x = rf,
                   xend = rf,
                   y = 1.799,
                   yend = 1.8),
               arrow = arrow(length = unit(2, "mm"),
                             type = "closed"),
               data = ~ dplyr::filter(., est + 1.96 * se > 1.8)) +
  geom_segment(aes(x = rf,
                   xend = rf,
                   y = 0.90001,
                   yend = 0.9),
               arrow = arrow(length = unit(2, "mm"),
                             type = "closed"),
               data = ~ dplyr::filter(., est - 1.96 * se < 0.9))
```
